---
title: "penguin flipper length ANOVA"
author: "Max Russer"
date: "November 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in the penguins.csv data

Packages: tidyverse, car, onewaytests

```{r message = FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")
```

Our question: Is there a significant difference in mean flipper lengths for female Chinstrap, Gentoo, and Adelie Penguins near Palmer Arhcipelago?

Maybe I can use a one-way ANOVA...
-Independent observations
- Equal variance (Levene's test)
- Normality (histogram/qqplots)

```{r}

counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>%
  mutate (
    common_name = 
      case_when(
        species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap", species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo", species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie"
      )
  ) %>%
  select(-species) %>%
  filter(sex == "FEMALE")

```

Exploratory graphs and a Levene's test for equal variances

```{r}

hist <- ggplot(penguins2, aes(x = flipper_length)) +
  geom_histogram(aes(fill = common_name)) +
  facet_wrap(~common_name, scale = "free")

hist

geom <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq(aes(color = common_name)) +
  facet_wrap(~common_name, scale = "free")

geom

# Based on qq-plots and histograms, observations for each group appear normally distributed. Also, n > 30 for all groups, so means will be normally distributed by CLT.

```

Evaluate variances and perform a Levene's Test for equal variances

```{r}

# If the largest sample variances is < 4x greater than the smallest sample variance , then these are "close enough" to consider variances equal

variances <- penguins2 %>%
  group_by(common_name) %>%
  summarize(
    variance = var(flipper_length)
  )

variances # The largest variance (33.1) is definitely not more than 4x greater than the smallest variance (15.2)

# Levene's Test
# LeveneTest function in the 'car' package

levene_penguin <- leveneTest(flipper_length ~ common_name, data = penguins2)

levene_penguin

# HO: No difference in variances (variances are equal)
# HA: Variances are NOT equal

# There is no significanct difference in variances across species. 
```

One-way ANOVA for penguin flipper length by species (common_name)


```{r}

# HO: Means across all groups are equal
# HA: At least two group means differ significantly

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# Post -hoc testing by Tukey's HSD
# HO: No significant difference in pairwise means
# HA: There is a significant difference in pairwise means

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

# aov.test function in the 'onwaytests' packages does allow for inline references

```

Mean flipper lengths of female penguins differed significantly for all species studied (one-way anova, F(df) = statistic, p < 0.001, $\alpha$ = 0.05 with post-hoc Tukey's HSD, $\alpha$ = 0.05)
